\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% Raphael and Pablo
% Clarisa Sánchez, Mathias Unberath, Tobias Nef (committee)
% Gonzalo & Luis + David
% Friends at AIMI
% Friends in Bern
% Friends in Spain
% Family
% Valentina

The decision to pursue a PhD came from the personal desire to contribute to the society that shaped my education and to add a piece to the immense ocean that human knowledge is. I hoped this journey would teach me humility, reminding me that what I don’t know far outweighs what I do. As I approach the final words of this dissertation, I can confidently say that those beliefs have not only held true, but deepened with time. Much has changed over the years, and I have learned a lot in this process, yet the initial spark that inspired me to work for the greater good has only become stronger. A PhD is filled with moments of triumph and challenge, and through both, I have leaned on the support of those around me. To each of you, I offer my sincere gratitude. This work would not have been possible without you.

Thanks to my supervisor, Raphael Sznitman, for allowing me to pursue a PhD in his group. I began this journey during a time of global uncertainty, but his flexibility, guidance, and unwavering support made it possible for me to return home when I needed it most. Alongside Pablo Márquez Neila, they became a source of inspiration with endless patience, always keen to engage with the ideas I proposed and uncover the subtleties I had overlooked. They transformed rough rocks into polished crystals that now deserve a place in this dissertation. For that, I sincerely thank them both.

Four years contain four summers, two of which I spent mentally (although not physically) away from the PhD. I want to thank Mahdi Azizian and Sean Huver for the trust they deposited in me during these summers, taking me as an intern and showing me the future of healthcare. Thanks also to Clarisa Sánchez, Mathias Unberath, and Tobias Nef for serving on my committee and to Stella\textsuperscript{\textcopyright} for serving me the food.

Our lives are shaped not only by the people around us but also by the ideas and voices we invite in. Countless podcasts have been my constant companions throughout this journey, and I cannot leave without acknowledging the impact that their authors have had on me. \textit{Lo que tú digas}, \textit{Nadie sabe nada}, \textit{Lex Fridman} or \textit{La Contracrónica}, among others, have been part of my daily routine, and for the knowledge and inspiration they have provided, I am deeply grateful.

My journey into AI started in January 2018 when Gonzalo Izaguirre drew my attention to a documentary on AlphaGo. Little did he know that this seemingly little nudge would transform my career. I am thankful for that and also for your invaluable help. Likewise, my mind goes to Luis Esteban, who was the first to know about my idea to pursue a PhD during a bus ride through Manchester in December 2019. Thank you for offering me a place in Madrid whenever I needed it and always being there.

I am grateful to everyone who has made this time as fun as it has been. I want to thank everybody in AIMI lab: Mathias, the first anchor when new members arrived and always happy to help. My fellow PhD students Lars, Lukas, Sergio, Fei, Paulo, and the later arrived Moritz with whom we spent most of the time, shared a room and a whiteboard that saw all our ideas. Marta, María, and Prisca from the Cool Table, and Davide for the fun during trips and skiing. In no particular order, my thanks also go to Alain, Heather, Lucas, Michel, Negin, Theo, Tea, other chess players (and non chess players) on the B-floor, and the OOC group on the E-floor.

Life isn’t just about work, and I’ve been fortunate to be surrounded by incredible people along the way. To my friends in Bern, Zaragoza, Madrid, and elsewhere, thank you for making this journey so much more enjoyable. Whether it was running, reading, discussing finance, or simply spending time together whenever I could visit, your presence has meant the world to me.

Pero nada de esto habría sido posible sin el apoyo incondicional de mi familia. Desde siempre, mis padres me han impulsado a seguir mis intereses, abriéndome las puertas al mundo de la tecnología y brindándome un entorno seguro en el que desarrollarme. Gracias.\sidenote{But none of this would have been possible without the unconditional support of my family. My parents have always encouraged me to follow my interests, opening the doors to the world of technology and giving me a safe environment in which to develop. Thank you.} 

Finally, I want to express my gratitude to Valentina. Your impact has extended far beyond the pages of this thesis.