\graphicspath{{ch4_fullweak/}{Figures/}}
% \graphicspath{{Figures/}}

\chapter[Full or Weak annotations?]{Full or Weak annotations? An adaptive strategy for budget-constrained annotation campaigns}
\label{chapter:fullweak}

\sidechaptersummary{Optimal budget allocation for image labeling, Future segmentation performance prediction,  Usage of Gaussian Processes as surrogate models}

\subsubsection{Synopsis}Annotating new datasets for machine learning tasks is tedious, time-consuming, and costly. For segmentation applications, the burden is particularly high as manual delineations of relevant image content are often extremely expensive or can only be done by experts with domain-specific knowledge. Thanks to developments in transfer learning and training with weak supervision, segmentation models can now also greatly benefit from annotations of different kinds. However, for any new domain application looking to use weak supervision, the dataset builder still needs to define a strategy to distribute full segmentation and other weak annotations. Doing so is challenging, however, as it is a priori unknown how to distribute an annotation budget for a given new dataset. To this end, in this chapter, we propose an approach to determine annotation strategies for segmentation datasets, whereby estimating what proportion of segmentation and classification annotations should be collected given a fixed budget. Our method sequentially determines proportions of segmentation and classification annotations to collect for budget-fractions by modeling the expected improvement of the final segmentation model. In our experiments, we show that our approach yields annotations that perform very close to the optimal for a number of different annotation budgets and datasets.

\subsubsection{Publication}This chapter is based on a publication\sideauthorcite{Tejero_2023_CVPR}, and its contents have been modified slightly to be more consistent with the rest of the thesis. More specifically, \Cref{tab:configurations_model,tab:configurations_gp} have been added to the main text from the supplementary material, as well as the ground truth approximation methodology in \Cref{sec:experiments_fullweak}.

\subsubsection{Author contributions}The work in this chapter was done in collaboration with the University Hospital Bern. The contributing authors were Pablo Márquez Neila, Sebastian Wolf, Martin Zinkernagel, and Raphael Sznitman. My contribution was the co-development and implementation of the method; conception and evaluation of the experiments and, together with the co-authors, writing of the paper.

\input{01_introduction}
\input{02_related}
\input{03_method}
\input{04_experiments}
\input{05_results}
\input{06_conclusion}